---
title: "R Programminfg for Research Book Summary"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##Setting the working director
setwd("")
getwd()
list.files()
## Chapter 2:  Entering and cleaning data
---
**Most ﬂat ﬁles come in two general categories:**
  1. Fixed width ﬁles : are ﬁles where a column always has the same width, for all the rows in the column.
  2. Delimited ﬁles:
    - “.csv”: Comma-separated values
    - “.tab,” “.tsv”: Tab-separated values
    - Other possible delimiters: colon, semicolon, pipe (“|”)
---
##Reading files from remote datasets
```
 url <- paste0("https://raw.githubusercontent.com/cmrivers/", "ebola/master/country_timeseries.csv")
 # ebola <- read_csv(file = url)
 ## slice(.data = (select(.data = ebola, 1:3)), 1:3)
 
 ```
---
##To build a logical expression to use in filter\you’ll need to know some of R’s logical operators.\ Some of the most commonly used ones are:

```{r }
library(knitr)
library(kableExtra)
library(tidyverse)

logical_operators <- data.frame(Operator=c('==', 
                                           '!=', 
                                           '%in% ',
                                           'is.na()', 
                                           '!is.na()', 
                                           '&',
                                           '|' ),
                 equals= c('equals', 
                           'does not equals', 
                           'is in category',
                           'is missing', 
                           'is not missing', 
                           'and',
                           'or' ),
                 Example = c('category == "Acting"', 
                             'category != "Comedy"', 
                             '%in% c("Academic", "Science")',
                             'is.na(job)', '!is.na(job)', 
                             'year == 2015 & category == "Academic"',
                             'year == 2015 | category == "Academic"' ))
knitr::kable(head(logical_operators[, 1:3]), "simple")
```

## Generic code for renaming variables
*rename(dataframe,
  new_column_name_1 = old_column_name_1,
  new_column_name_2 = old_column_name_2)*
  
  
```{r}
#Loading the required libraries
library("readr")
library("dplyr")


#Loading the required data


daily_show <- read_csv(file = "daily_show_guests.csv",show_col_types = FALSE)
head(daily_show)
#See the column names
colnames(daily_show)


#Renaming the column names
daily_show<- as_tibble(select(daily_show, 1:5))

daily_show <-rename(daily_show,
                    year = YEAR,
                    job = GoogleKnowlege_Occupation,
                    date = Show,
                    category = Group,
                    guest_name = Raw_Guest_List)

head(x = daily_show, 3)


#Selecting columns
select(.data = daily_show, job, date, category, guest_name)
#Slicing can extract certain rows based on their position in the dataframe
slice(.data = daily_show, 1:3)
#To extract two random rows from the dataframe
sample_n(tbl = daily_show, size = 2)

#arrange(.data = dataframe, column_to_order_by)
daily_show <- arrange(.data = daily_show, category)
slice(.data = daily_show, 1:3)
#Arranging the daatasests in dsescending orders
daily_show <- arrange(.data = daily_show,
                      desc(x = year))
slice(.data = daily_show, 1:2)

#Filtering certain rows
scientists <- filter(.data = daily_show,
                     category == "Science")
head(x = scientists)

```
#Chaining all the operations through pipe operator
daily_show %>% rename(job = GoogleKnowlege_Occupation,date = Show,category = Group,guest_name = Raw_Guest_List) %>%
                      select(-YEAR) %>%
                      mutate(job = str_to_lower(job)) %>%
                      filter(category == "Science")
##Dates in R

knitr::kable(data,
 caption = "Table with kable")
 
| Right | Right|
|-------:|:------|
|  **Abbreviation** | **Meaning**| 
| %m | Month as a number (e.g., 1, 05) | 
| %B  |  Full month name (e.g., August)|
|%b|Abbreviated month name (e.g., Aug)| 
| %y | Two-digit year (e.g., 99)|
| %Y|Four-digit year (e.g., 1999)|

##Working with dates

knitr::kable(data,
 caption = "Abbreviation Date")
 
| Right | Right|
|-------:|:------|
|  Current format of date | format = | 
| 10/23/2008 | “%m/%d%Y” | 
| 08-10-23  |  “%y-%m-%d”|
| Oct. 23 2008 |“%b. %d %Y”| 
| October 23, 2008  |  “%B %d, %Y”|

```{r , echo=FALSE}
#Changing the column names
library(package = "stringr")

mutate(.data = daily_show,
       job = str_to_lower(string = job))

#To convert a vector to the Date class, you can use as.Date():
daily_show <- mutate(daily_show,
                     date = as.Date(date, format = "%m/%d/%y"))
head(daily_show, 3)
class(daily_show$date)
#calculate the range of dates
range(daily_show$date)
#Calculate total number of days datsets covers
diff(range(daily_show$date))
```

```{r}
# Use `read_tsv` to read this file. Because the first line
# of the file is *not* the column names, you need to specify what the column
# names should be with the `col_names` parameter.

#List all the available data in the directory
list.files("./data")

library(readr)
ca_measles <- read_tsv(file = "data/measles_data/02-09-2015.txt", col_names = c("city", "count"))
head(x = ca_measles)


# You'll need the `readxl` package to read in the Excel file. Load that.
library(package = "readxl")
# Use the `read_excel` function to read in the file.
icd10 <- read_excel(path = "data/icd-10.xls")
head(x = icd10)


# You'll need the `haven` function to read in the SAS file. Load that.
library(package = "haven")
# Use the `read_sas` function to read in this file.
icu <- read_sas(data_file = "data/icu.sas7bdat")
library(package = "dplyr")
slice(.data = select(.data = icu, 1:5), 1:5)

```





```{r}
# Copy this code to an R script and add comments describing what each line is doing
# Install any packages that the code loads but that you don't have.
library(package = "haven")
library(package = "forcats")
library(package = "stringr")

# Reading SAS data and recoding factors
library(package = "haven")
icu <- read_sas(data_file = "data/icu.sas7bdat") %>%
  select(ID, AGE, GENDER) %>%
  rename(id = ID,
         age = AGE,
         gender = GENDER) %>%
  mutate(gender = as_factor(x = gender),
         gender = fct_recode(.f = gender,
                             Male = "0",
                             Female = "1"),
         id = str_c(id)) %>%
  arrange(age) %>%
  slice(1:10)

icu

ebola_liberia <- read_csv(file = "data/country_timeseries.csv") %>%
  select(Date, Cases_Liberia, Deaths_Liberia) %>%
  rename(date = Date,
         cases = Cases_Liberia,
         deaths = Deaths_Liberia) %>%
  mutate(ratio = deaths / cases) %>%
  filter(!is.na(x = cases))
head(x = ebola_liberia)
```

##CHAPTER 3 : Exploring data

*-range(),min(), max(),mean(), median(),sd() ,table() ,cor(),summary(),
-Range (minimum and maximum) of vector, Minimum or maximum of vector, Mean or median of vector, 
Standard deviation of vector, Number of observations per level for a factor vector,
Determine correlation(s) between two or more vectors,
Summary statistics depends on class*
**all of these functions takes the *na.rm = TRUE* 

All of these functions take, as the main argument, the vector or vectors for which
you want the statistic. If there are missing values in the vector, you’ll typically need
to add an argument to say what to do with the missing values. The parameter name
for this varies by function, but for many of these functions it’s na.rm = TRUE or
use="complete.obs".
##GGPLOT2 
Code Description
‘x‘ Position on x-axis
‘y‘ Position on y-axis
‘shape‘ Shape
‘color‘ Color of border of elements
‘ﬁll‘ Color of inside of elements
‘size‘ Size
‘alpha‘ Transparency (1: opaque; 0: transparent)
‘linetype‘ Type of line (e.g., solid, dashed)

There are also a number of elements that you can add onto a ggplot object using +.
A few that are used very frequently are:
Element Description
‘ggtitle‘ Plot title
‘xlab‘, ‘ylab‘ x- and y-axis labels
‘xlim‘, ‘ylim‘ Limits of x- and y-axis
```{r}
library(faraway)
library(dplyr)
library(ggplot2)

data(nepali)

#Each observation is a single measurement for a child; there can be multiple observations per child. Let's subset out child id, sex, weight, height, and age, and then limit to each child’s ﬁrst measurement.

nepali <- nepali %>%
  # Subset to certain columns
  select(id, sex, wt, ht, age) %>%
  # Convert id and sex to factors
  mutate(id = factor(id),
         sex = factor(sex, levels = c(1, 2),
                      labels = c("Male", "Female"))) %>%
  # Limit to first obs. per child
  distinct(id, .keep_all = TRUE)
head(nepali)


#For geom_histogram(), the main aesthetic is x, the (numeric) vector for which you want to create a histogram:

ggplot(nepali, aes(x = ht)) +
  geom_histogram(fill = "lightblue", color = "black", bins = 40) +
  ggtitle("Height of children") +
  xlab("Height (cm)") + xlim(c(0, 120))

ggplot(nepali,aes(x = ht, y = wt, color = sex)) +
  geom_point(color = "blue", size = 0.5) +
  ggtitle("Weight versus Height") +
  xlab("Height (cm)") + ylab("Weight (kg)")

#boxplots by a factor. In this case, you’ll need to include two aesthetics (x and y) when you initialize the ggplot object.

ggplot(nepali, aes(x = sex, y = ht, group = sex)) +
  geom_boxplot() +
  xlab("Sex")+ ylab("Height (cm)")

#For example, you can use the ggpairs function from the GGally package to plot all pairs of scatterplots for several variables. Notice how this output shows continuous and binary variables diﬀerently.

library(GGally)
ggpairs(nepali[, c("sex", "wt", "ht", "age")])
cor(nepali[ , c("wt", "ht", "age")], use = "complete.obs")

#ummaries by group using dplyr (e.g., mean weight by sex), use group_by before running summarize:

nepali %>%
  group_by(sex) %>%
  summarize(mean_wt = mean(wt, na.rm = TRUE),
            n_children =n(),
            first_id = first(id))

nepali %>%
  mutate(is_male = sex == "Male") %>%
  group_by(is_male) %>%
  summarize(n_children = n())
```

```{r}
library(package = "lubridate")

class(x = daily_show$date)

daily_show1 <- mutate(.data = daily_show,
                      date = ymd(date))
head(x = daily_show, n = 3)
range(daily_show$date)
diff(x = range(daily_show$date))
```
The lubridatepackage also includes functions to pull out certain elements of a date,
including:
• wday
• mday
• yday
• month
• quarter
• year

```{r}
mutate(.data = daily_show1,
show_day = wday(x = date, label = TRUE)) %>%
select(date, show_day, guest_name) %>%
slice(1:5)
```

```{r}
library(faraway)
data("worldcup")
worldcup %>%
  group_by(Position) %>%
  summarize(max_saves = max(Saves))

worldcup %>%
  filter(Position == "Forward") %>%
  group_by(Team) %>%
  summarize(n_forwards = n(),
            total_forward_shots = sum(Shots)) %>%
  arrange(desc(total_forward_shots))

worldcup %>%
  filter(Position == "Defender") %>%
  top_n(n = 1, wt = Tackles)

brazil_players <- worldcup %>%
  filter(Team == "Brazil" & Position != "Goalkeeper")
head(brazil_players)

ggplot(worldcup, mapping = aes(x = Time, y = Passes, color = Position)) +
  geom_point() +
  geom_rug()

worldcup %>%
  rownames_to_column(var = "Name") %>%
  filter(Team %in% c("Spain", "Netherlands", "Germany", "Uruguay")) %>%
  ggplot() +
  geom_point(aes(x = Shots, y = Tackles, color = Position, shape = Team)) +
  geom_text(mapping = aes(x = Shots, y = Tackles,
                          color = Position, label = Name),
            size = 2.5)

worldcup %>%
  mutate(top_4 = Team %in% c("Spain", "Netherlands", "Germany", "Uruguay")) %>%
  ggplot() +
  geom_point(aes(x = Time, y = Passes, color = top_4))

top_teams <- worldcup %>%
  filter(Team %in% c("Spain", "Netherlands", "Germany", "Uruguay"))

ggplot(data = top_teams) +
  geom_histogram(aes(x = Time, fill = Team))

ggplot(top_teams, aes(x = Team, y = Shots)) +
  geom_boxplot() +
  ggtitle("Shots per player in World Cup 2010")

ggplot(data = top_teams) +
  geom_histogram(aes(x = Time, fill = Team), binwidth = 90)
```



##Chapter 4 : Reporting data results #1

```{r}
ggplot(worldcup, aes(x = Time, y = Passes,
                     color = Position, size = Shots)) +
  geom_point(alpha = 0.5) +
  scale_x_continuous(name = "Time played (minutes)",
                     breaks = 90 * c(2, 4, 6),
                     minor_breaks = 90 * c(1, 3, 5))

ggplot(chic_july, aes(x = date, y = death)) +
  geom_line()

library(RColorBrewer)
display.brewer.pal(name = "Set1", n = 8)
display.brewer.pal(name = "PRGn", n = 8)
display.brewer.pal(name = "PuBuGn", n = 8)

a <- ggplot(data.frame(x = 1:5, y = rnorm(5),
                       group = letters[1:5]),
            aes(x = x, y = y, color = group)) +
  geom_point()
b <- a + scale_color_brewer(palette = "Set1")
c <- a + scale_color_brewer(palette = "Pastel2") +
  theme_dark()
grid.arrange(a, b, c, ncol = 3)

ggplot(worldcup, aes(x = Time, y = Passes,
                     color = Position, size = Shots)) +
  geom_point(alpha = 0.5) +
  scale_color_brewer(palette = "Dark2",
                     name = "Player position")

#You can also set colors manually:

ggplot(worldcup, aes(x = Time, y = Passes,color = Position, size = Shots)) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("blue", "red","darkgreen", "darkgray"))

#To add a text label with just the player with the most shots, you’ll want to create a new dataframe with just the top player. You can use the top_n function to do that (the wtoption is specifying that we want the top player in terms of values in the Shots column):

top_player <- worldcup %>%
  top_n(n = 4, wt = Shots) %>%
  mutate(label = paste0(Player, ", ", Team))

worldcup %>%
  mutate(top_4 = Team %in% c("Netherlands", "Uruguay", "Spain", "Germany")) %>%
  ggplot(aes(x = Time, y = Shots, color = top_4)) +
  geom_vline(xintercept = 90 * 3, color = "gray", linetype = 2) +
  geom_point(alpha = 0.5, size = 1) +
  geom_text(data = top_player, aes(label = label, color = NULL), hjust = 1.1, vjust = 0.4) +
  labs(x = "Time played in World Cup (minutes)", color = "Team's final\nranking") +
  theme_few()
```



