---
title: "R Programminfg for Research Book Summary"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Chapter 2:  Entering and cleaning data

Most ﬂat ﬁles come in two general categories:
  1. Fixed width ﬁles : are ﬁles where a column always has the same width, for all the rows in the column.
  2. Delimited ﬁles:
    • “.csv”: Comma-separated values
    • “.tab,” “.tsv”: Tab-separated values
    • Other possible delimiters: colon, semicolon, pipe (“|”)

#Reading files from remote datasets
 url <- paste0("https://raw.githubusercontent.com/cmrivers/", "ebola/master/country_timeseries.csv")
 # ebola <- read_csv(file = url)
 ## slice(.data = (select(.data = ebola, 1:3)), 1:3)

#To build a logical expression to use in filter, you’ll need to know some of R’s logical operators. Some of the most commonly used ones are:

Operator Meaning Example
== equals                         category == "Acting"
!= does not equal             category != "Comedy"
%in% is in category             %in% c("Academic", "Science")
is.na() is missing                is.na(job)
!is.na() is not missing         !is.na(job)
& and                             year == 2015 & category == "Academic"
| or                            year == 2015 | category == "Academic"
```{r}
#Loading the required libraries
library("dplyr")


#Loading the required data
daily_show <- read_csv(file = "D:/R_Programming/Github/SpatialDataWithR/data/daily_show_guests.csv", skip = 4)
head(daily_show)
#See the column names
colnames(daily_show)


#Renaming the column names
daily_show <- rename(.data = daily_show, 
                     year = YEAR,
                     job = GoogleKnowlege_Occupation,
                     date = Show,
                     category = Group,
                     guest_name = Raw_Guest_List)
head(x = daily_show, 3)


#Selecting columns
select(.data = daily_show, job, date, category, guest_name)
#Slicing can extract certain rows based on their position in the dataframe
slice(.data = daily_show, 1:3)
#To extract two random rows from the dataframe
sample_n(tbl = daily_show, size = 2)

#arrange(.data = dataframe, column_to_order_by)
daily_show <- arrange(.data = daily_show, category)
slice(.data = daily_show, 1:3)
#Arranging the daatasests in dsescending orders
daily_show <- arrange(.data = daily_show,
                      desc(x = category))
slice(.data = daily_show, 1:2)

#Filtering certain rows
scientists <- filter(.data = daily_show,
                     category == "Science")
head(x = scientists)

#Changing the column names
library(package = "stringr")

mutate(.data = daily_show,
       job = str_to_lower(string = job))


```
#Chaining all the operations through pipe operator
daily_show %>% rename(job = GoogleKnowlege_Occupation,
                      date = Show,
                      category = Group,
                      guest_name = Raw_Guest_List) %>%
 select(-YEAR) %>%
 mutate(job = str_to_lower(job)) %>%
 filter(category == "Science")
 
 

